\documentclass[10pt]{article}

\setlength{\parindent}{0eM}

\setlength{\parskip}{0ex}

\pagestyle{empty}

\usepackage{amssymb,amsmath,amsthm}

\title{MATH 351 Assignment 2 Exercise 2}
\author{\textsc{\textit{Steven Sharp}}}
\date{}

\begin{document}

\maketitle

\thispagestyle{empty}

\begin{center}
\rule{\textwidth}{0.5pt}
\end{center}

\setlength{\parindent}{1.5eM} I'm not sure if these are supposed to be just typesetting errors, or also grammatical errors, so I've included both in my list:

\setlength{\parindent}{0eM} 1. There is no period after the matrix $A$ to complete the sentence.

2. The \verb!\implies! command should not be used in place of the word implies outside of the field of formal logic.

3. $-2$ is not written in math mode, so it appears as -2.

4. In the centered equation, the \verb!\left! and \verb!\right! commands are omitted, so the parentheses do not scale with the integral.

5. In all integrals, the \verb!\,! command is omitted before $dx$ so that the spacing is too tight.

6. Equations with multiple equalities should appear on multiple lines.

7. The comma after the centered equation is unnecessary as it fragments the sentence.

8. The determinant operator should be displayed as $\det A$ or $\det(A)$, not $det(A)$.

9. In the first line, $A$ is erroneously displayed as A since it is left in text mode.

10. The second line is not indented, presumably because a line break was not included after the matrix equation in the \LaTeX{} code.

\begin{center}
\rule{\textwidth}{0.5pt}
\end{center}

\end{document}
